---
title: "Cepheids Period-Luminosity Relationship"
output: html_document
---
This is a recreation of the analysis done by Henrietta Leavitt on 25 Cepheid variables observed in the Small Magellanic Cloud.  From this analysis, the Period-Luminosity relationship for Cepheid variables was derived, and applied as a standard candle to determine interstellar distances.

Initial data from following paper:
Title: Periods of 25 Variable Stars in the Small Magellanic Cloud.
Authors: Leavitt, H. S. & Pickering, E. C.
Journal: Harvard College Observatory Circular, vol. 173, pp.1-3
Bibliographic Code: 1912HarCi.173....1L

Link: http://articles.adsabs.harvard.edu/full/1912HarCi.173....1L

Load the data

```{r}
cepheidData<-read.csv("data/cepheids.csv")
```

Preliminary plot of minimum

```{r}
plot(cepheidData$Min~log10(cepheidData$Period), ylim=rev(range(cepheidData$Min)))
```

Preliminary plot of maximum

```{r}
plot(cepheidData$Max~log10(cepheidData$Period), ylim=rev(range(cepheidData$Max)))
```

Preliminary fits
```{r}
modelMax<-lm(Max~log10(Period),data=cepheidData)
modelMin<-lm(Min~log10(Period),data=cepheidData)
```

Distance to SMC is 197 +/- 9 kly (info from Wikipedia),
or 61 +/1 kiloparsecs

Calculations are in ky, so we need a conversion from parsec to ly

```{r}
parsec <- 3.26156
```

Absolute-apparent magnitude formula
```{r}
absMag <- function(appMag,distObj=10*parsec) {
  return (appMag-5*(log10(distObj/parsec)-1))
}
```

Apparent-absolute magnitude formula
```{r}
appMag <- function(absMag,distObj=10*parsec) {
  return (absMag+5*(log10(distObj/parsec)-1))
}
```

Calculating distance from absolute and apparent magnitudes
```{r}
distObj <- function(absMag,appMag) {
  return (parsec*(10^(((appMag-absMag)/5)+1)))
}
```

Converting data to long format

```{r}
distance<-200
rawCeph <- read.csv("cepheids.csv")
cleanCeph<-gather(rawCeph,magType,appMag,Min:Max)
cleanCeph$absMag <- absMag(cleanCeph$appMag,distance)
```

and plotting in ggplot
```{r}
p <- ggplot(cleanCeph,aes(log(Period), absMag))
p <- p + geom_point(aes(color=magType))
p <- p + scale_y_reverse()
p
```

doing fits on long data

```{r}
cephFitsApp <- fitCephData(cleanCeph)
```